### 2017-02-14 paper/oopsla17-submission/introduction.tex
init paper

\section{Introduction}
\label{sec:introduction}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: