### 2017-03-28 paper/oopsla17-submission/discussion.tex
stub for discussion

\section{Discussion}
\label{sec:discussion}

\paragraph{Additional Features}
\begin{itemize}
\item def-use chains
\item connect function and arguments in application (right now each only know about the application itself)
\item ??
\end{itemize}

\paragraph{User-Defined Types}
\begin{itemize}
\item Hard to support in general, as each data and type constructor gets
  its own set of syntactic features.
\item Attempt to mitigate by adding generic ``is-datacon'' and
  ``is-match'' feature, unclear how much it will help
\end{itemize}

\paragraph{Independent vs Joint Predictions}
We treat each sub-expression as if it exists in a vacuum, in reality
making a positive prediction for one expression may make another
prediction much less likely \citep{Raychev2015-jg}. We compensate
somewhat by ranking predictions by ``confidence'', but would be
interesting to see how structured learning would perform.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: