### 2017-04-13 paper/oopsla17-submission/gallery.tex

\subsection{Explaining Predictions}
\label{sec:qualitative}

Next, we present a \emph{qualitative} evaluation that compares the
predictions made by our classifiers with those of \sherrloc.
%
In particular, we demonstrate, with a series of example programs from
our student dataset, how our classifiers are able to use past student
mistakes to make more accurate predictions of future fixes.
%
We also take this opportunity to examine some of the specific features
our classifiers use to assign blame.
%
For each example, we provide
%
(1) the code;
%
(2) \hlSherrloc{\sherrloc's} prediction;
%
(3) our \hlTree{\dectree's} prediction; and
%
(4) an \emph{explanation} of why our classifier made its prediction, in
terms of the features used and their values.
%
We choose the \dectree classifier for this section as its model is more
easily interpreted than the MLP.\@
%
We also exclude the \ExprSize feature from the model used in this
section, as it makes the predictions harder to motivate, and as we saw
in \autoref{sec:feature-utility} it does not appear to contribute
significantly to the model's accuracy.

We explain the predictions by analyzing the paths induced in
the decision tree by the features of the input programs.
\ES{TODO: EXPOUND!!}
%
% We will also attempt, for each program, to explain \emph{why} the
% Decision Tree made its prediction, by analyzing the paths induced
% by the programs.

\subsubsection{Failed Predictions}
\label{sec:failed-predictions}
We begin with a few programs where our classifier fails to
make the correct prediction.
%
For these programs we will additionally \hlFix{highlight} the
correct blame location.

\paragraph{Constructing a List of Duplicates}
% data/sp14/0219.ml
Our first program is a simple recursive function |clone| that takes an
item |x| and a count |n|, and produces a list containing |n| copies of
|x|.
%
% \begin{ecode}
%   let rec clone x n =
%     let loop acc n =
%       if n <= 0 then
%         acc
%       else
%         (*@\colorbox{red!50}{clone}@*) ==([==_=x=_==] @ acc)== (n - 1) in
%     loop [] n
% \end{ecode}
\begin{ecode}
  let rec clone x n =
    let loop acc n =
      if n <= 0 then
        acc
      else
        (*@\hlFix{clone} \hlSherrloc{([\hlTree{x}] @\ acc)}@*) (n - 1) in
    loop [] n
\end{ecode}
% \ES{TODO: our 2nd prediction matches \sherrloc and \ocaml (occurs check), correct fix is to replace recursive call to clone with loop}
%
The student has defined a helper function |loop| with an accumulator
|acc|, likely meant to call itself tail-recursively.
%
Unfortunately, she has called the top-level function |clone| rather than
|loop| in the |else| branch, this induces a cyclic constraint |'a = 'a list|
for the |x| argument to |clone|.

Our classifier incorrectly predicts that the use of |x| in the recursive
call is the most likely source of the error.
%
Our second prediction coincides with \sherrloc (and \ocaml), blaming the
the first argument to |clone|.
%
This is also incorrect, but may be more helpful than our first
prediction --- if our student decides that she has certainly provided
the correct \emph{argument}, an alternative explanation is that
perhaps she has called the wrong \emph{function}.

We confess that both of these predictions are difficult to explain by
examining the induced paths.
%
In particular, neither prediction references a property of the current
expression, only its context, which is surprising.
%
Much clearer is why we fail to blame the occurrence of |clone|, the two
enabled features on the path are:
%
(1) the parent is an application; and
%
(2) |clone| has a function type.
%
The model appears to have learned that programmers typically call the
correct function.

\paragraph{Currying Considered Harmful?}
% data/sp14/1204.ml
Our next example is another ill-fated attempt at |clone|.
%
\begin{ecode}
  let rec clone x n =
    let rec loop (*@\hlFix{x n acc}@*) =
      if n < 0 then
        acc
      else
        (*@\hlSherrloc{loop} \hlTree{(x, (n - 1), (x :: acc))}@*) in
    loop (x, n, [])
\end{ecode}
%
The issue here is that \ocaml functions are \emph{curried} by default
--- \ie they take their arguments one at a time --- but our student has
called the inner |loop| with all three arguments in a tuple.
%
Many experienced functional programmers would choose to keep |loop|
curried and rewrite the calls, however our student decides instead to
\emph{uncurry} |loop|, making it take a tuple of arguments.
%
\sherrloc blames the recursive call to |loop| while our classifier
blames the tuple of arguments --- a reasonable suggestion, but not
the answer the student expected.

We fail to blame the definition of |loop| because it is defining a
function.
%
First, note that we represent |let f x y = e|\ \ as\ \ |let f = fun x -> fun y -> e|,
thus a change to the pattern |x| would be treated as a change to the outer
|fun| expression.
%
With this in mind, we can explain our failure to blame the definition of
|loop| (the outer |fun|) as follows:
%
(1) it has a function type;
%
(2) its child is a |fun|; and
%
(3) its parent is a |let|.
%
Thus it appears to the model that the outer |fun| is simply part of a
function definition, a common and innocuous phenomenon.

% \paragraph{Composing Functions}
% % data/sp14/3041.ml
% Next, let us consider the |pipe| function that composes a list of
% functions, \ie given a list of functions |[f;g;h]|, |pipe| produces the
% function |fun x -> h (g (f x))|.
% %
% \begin{ecode}
%   let pipe fs =
%     let f a x y = ==y== __(a y)__ in
%     let base x = x in
%     List.fold_left f base fs
% \end{ecode}
% %
% The error in our student's code is that she has applied |y| rather than
% |x| to the result of |a y|.
% %
% \sherrloc correctly blames the first occurrence of |y|, while our
% classifier (incorrectly) blames the application |a y| (\ocaml blames
% the occurrence of |base| on line 4).
% %
% \ES{this has the same explanation as the 1st clone above}

\subsubsection{Correct Predictions}
\label{sec:correct-predictions}
Next, we present a few indicative programs where our first prediction is
correct, and all of the other tools' top three predictions are
incorrect.

\paragraph{Extracting the Digits of an Integer}
% data/sp14/2176.ml
Consider first a simple recursive function |digitsOfInt| that extracts
the digits of an |int|.
%
\begin{ecode}
  let rec digitsOfInt n =
    if n <= 0 then
      []
    else
      [n mod 10] @ (*@\hlTree{[ \hlSherrloc{digitsOfInt (n / 10)} ]}@*)
\end{ecode}
%
Unfortunately, the student has decided to wrap the recursive call to
|digitsOfInt| with a list literal, even though |digitsOfInt| already
returns an |int list|.
%
Thus, the list literal is inferred to have type |int list list|, which
is incompatible with the |int list| on the left of the |@| (list append)
operator.
%
Both \sherrloc and the \ocaml compiler blame the recursive call for
returning a |int list| rather than |int|, but the recursive call is
correct!

As our \dectree correctly points out (with high confidence), the fault
lies with the list literal \emph{surrounding} the recursive call, remove
it and the type error disappears.
%
An examination of the path induced by the list literal reveals that our
\dectree is basing its decision on the fact that
%
(1) the expression is a list literal;
%
(2) the child expression is an application, whose return type mentions |int|; and
%
(3) the parent expression's type mentions |list|.
%
Interestingly, \dectree incorrectly predicts that the child application
should change as well, but it is less confident of this prediction and
ranks it below the correct blame assignment.

\paragraph{Padding a list}
% data/sp14/0442.ml
Our next program, |padZero|, is given two |int list|s as input, and must
left-pad the shorter one with enough zeros that the two output lists
have equal length.

The student first defines a helper function |clone|, as seen above.
%
\begin{ecode}
  let rec clone x n =
    if n <= 0 then
      []
    else
      x :: clone x (n - 1)
\end{ecode}
%
Then she defines |padZero| with a simple branch that determines which
list is shorter, followed by a |clone| for the appropriate number of
zeros.
%
\lstset{firstnumber=last}
\begin{ecode}
  let padZero l1 l2 =
    let n = List.length l1 - List.length l2 in
    if n < 0 then
      (clone 0 ((-1) * n) @ l2, l2)
    else
      (l1, (*@\hlTree{\hlSherrloc{clone 0 n} :: l2}@*))
\end{ecode}
\lstset{firstnumber=1}
%
Alas, our student has accidentally used the |::| operator rather than
the |@| operator in the |else| branch.
%
\sherrloc and \ocaml correctly determine that she cannot cons the
|int list| returned by |clone| onto |l2|, which is another |int list|,
but they decide to blame the call to |clone|, while our
\dectree correctly blames the |::| constructor.

Examining the path induced by the |::| constructor, we can see that our
\dectree is influenced by the fact that:
%
(1) the expression is a data constructor (not specifically |::|);
%
(2) the parent expression is a tuple; and
%
(3) the leftmost child is an application.
%
We note that first fact appears to be particularly significant; an
examination of the training samples that reach that decision node
reveals that before observing the \textsc{Is-Constructor} feature, the
classifier is slightly in favor of predicting ``blame'', but afterwards
it is heavily in favor of predicting ``blame''.
%
Many of the following decisions change the balance back towards ``no
blame'' if the ``true'' path is taken, but the |::| constructor always
takes the ``false'' path.
%
It would appear that our \dectree has learned that data constructors are
particularly suspicious, and is looking for exceptions to this general
rule.
%
% \ES{FIXME: this is probably super confusing, checking with Huma to see
%   if we can get a nice graph to illustrate what's happening..}

Our \dectree correctly predicts that the recursive call blamed by
\sherrloc should not be blamed; a similar examination suggests that the
crucial observation is that the recursive call's parent is a data
constructor application.
%

% ES: decision tree gets this one wrong, may want to find something else
% \begin{ecode}
%   let rec sepConcat sep sl =
%     match sl with
%     | [] -> ""
%     | h::t ->
%         let f a x = a ^ (sep ^ x) in
%         List.fold_left f h t

%   let stringOfList f l = sepConcat "; " __[ "["; ___=List.map f l=___; "]" ]__
% \end{ecode}

% \ES{I don't much like this example anymore..}
% \paragraph{Computing the Fixed Point of a Function}
% % data/sp14/1266.ml
% Finally, our students must write a |fixpoint| function that computes the
% fixed point of a given function |f|, starting from an initial value |b|.
% %
% As a hint, we first have them write a |wwhile| function that performs the
% functional equivalent of a \texttt{while}-loop, repeatedly passing a function's
% output back in until it receives a (boolean) signal to stop.
% %
% \begin{ecode}
%   let rec wwhile (f, b) =
%     match f b with
%     | (x, false) -> x
%     | (x, true)  -> wwhile (f, x)
% \end{ecode}
% \lstset{firstnumber=last}
% %
% The |fixpoint| function can then be written as a clever instantiation of
% the arguments to |wwhile|.
% %
% \begin{ecode}
%   let fixpoint (f, b) =
%     let g = __let bb = f b in ___=(bb, (bb = b))=_ in
%     wwhile (g, b)
% \end{ecode}
% \lstset{firstnumber=1}
% %
% Sadly, our student has forgotten that |g| should itself be a function.
% %
% As a result, she passes a pair of a \emph{pair} and a starting value to
% |wwhile|, rather than a pair of a \emph{function} and a starting value.
% %
% \sherrloc deduces that the call to |wwhile| is likely correct (\ocaml
% actually blames the use of |g|), but identifies the construction of the
% pair inside |g| as the most likely culprit, while our \dectree correctly
% identifies the \emph{definition} of |g| as the source of the error.
% %
% Our \dectree appears to base its prediction, with complete confidence,
% on the fact that:
% %
% (1) the parent expression is a |let|;
% %
% (2) the right child is a tuple; and finally
% %
% (3) the current expression is a |let|.
% %

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: