### 2017-04-13 paper/oopsla17-submission/abstract.tex
asd

\begin{abstract}
  Localizing type errors is challenging in languages with global type
  inference, as the type checker must make assumptions about what the
  programmer intended to do.
  %

  We present a data-driven approach to type error localization that uses
  machine learning to assign blame based on previous
  ill-typed programs and their fixes.
  %
  First, we gather a large set of over 4,500 ill-typed student programs
  and their subsequent fixes.
  %
  Second, we frame type error localization as a supervised learning
  problem: to predict, for each expression in the
  program, whether it should be blamed.
  %
  We represent each expression by its syntactic class and type and those
  of its parent and children, associating blame with expressions that
  were changed in the human-fixed program.
  %
  Finally, we train and evaluate a series of off-the-shelf classifiers
  on our suite of programs,
  and find that they improve localization accuracy from a state-of-the-art
  56\% to 72\%.
\end{abstract}