### 2017-04-13 paper/oopsla17-submission/abstract.tex
asd

\begin{abstract}
Localizing type errors is challenging in
languages with global type inference, as
the type checker must make assumptions
about what the programmer intended to do.
%
We introduce \toolname, a \emph{data-driven}
approach to error localization based on
supervised learning.
%
\toolname works by using a large corpus
of training data --- pairs of ill-typed
programs and their ``fixed'' versions ---
to automatically \emph{learn a model}
of where the error is most likely.
%
When given a new ill-typed program,
\toolname executes the model to
generate a list of likely error
locations ranked by likelihood.
%
We evaluate \toolname by comparing its
precision against the state-of-the-art
on a set of over 4,500 ill-typed Ocaml
programs drawn from introductory
programming classes.
%
We show that when the top-ranked location
is considered, \toolname's data-driven
model is able to correctly predict
the exact sub-expression that needs to
be changed, 74\% of the time, which is
29 points higher than \ocaml, and
18 points higher than the state-of-the-art
\sherrloc tool.
%
Further, \toolname's accuracy surpasses
85\% when we consider the \emph{top-2}
locations and reaches 91\% if consider
the \emph{top-3}.
\end{abstract}