### 2017-04-13 paper/oopsla17-submission/conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

We have presented \toolname which
combining modern statistical methods
with domain-specific feature engineering
to open the door to a new data-driven
path to precise error localization.
%
\toolname
translates pairs of
blame-labeled ill-typed
programs into bags-of-abstracted-terms,
and then trains a classifier that
predicts the value of the
whether a feature vector gets
a blame label.
%
Given a new ill-typed program,
we extract its bag-of-extracted-terms
to compute the blame label for each
term and rank the predictions based
on the model's confidence, to select
the top 1-3 blame assignments to
present to the user.

We have evaluated our approach with
five different classifiers on a suite
of over 4,500 student programs.
We find that while machine
learning on its own performs slightly worse
than existing purely constraint based
approaches (\eg \ocaml, \sherrloc), once
we augment the data with feature abstraction
corresponding to the \emph{type error slice}~\cite{Tip2001-qp},
the resulting models models outperform
the state of the art by up to 18
percentage points.

As with other forms of machine learning,
a key concern is that of \emph{data-set bias}: are
the models learned by \toolname specific
to our data set, and would they fail on
\emph{other} \ocaml programs?
%
We address this concern in two ways.
%
First, we partition the data by year,
and show that models learned from one
year generalize to, \ie perform nearly
as well on the programs from the other
year.
%
Second, we argue that in our setting
this bias is \emph{feature} (and not
a bug): it allows \toolname to \emph{adapt}
to the kinds of errors that programmers
(specifically, novices who are in greatest
need of precise feedback) actually make
rather than hardwiring the biases of
compiler authors who, by dint of their
training and experience, may suffer from
blind spots with regards to such problems.
%
In this regard, we are particularly pleased
that our classifiers can be trained on a
modest amount of data, \eg a single course's
worth, and envision a future where each course
has a model of its students' errors.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: