### 2017-04-13 paper/oopsla17-submission/scratch.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Consider the |sumList| program in \autoref{fig:sumList},
written by a student in an undergraduate Programming Languages
course.
%
The program is meant to sum the integers in a list, but
the student has accidentally given the empty list as the
base case, rather than 0.
%
The \ocaml compiler collects typing constraints as it
traverses the program, and reports an error the moment
it finds an inconsistent constraint.
%
In this case it blames the recursive call to |sumList|,
complaining that |sumList| returns a |list| where an |int|
was expected by the |+| operator.
%
Note that this \emph{blame} is inconsistent with the
programmer's intention and may not be helpful debugging
information to a novice.

It may appear obvious to the reader that
|[]| is the correct expression to blame,
but how is the type checker to know that?

Indeed, both the venerable
Ocaml compiler as well as
recent techniques~\citep{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr}.
fail to distinguish between
the two underlined expressions
in \autoref{fig:sumList};
it would be equally valid
to blame \emph{either}
of them alone.
%
The |[]| on line 3 could be changed to |0|,
or the |+| on line 4 could be changed to
either |@| (list append) or |::|, all of
which would give type-correct programs.
%
Thus, these state-of-the-art techniques
are forced to either blame \emph{both}
locations, or choose one \emph{arbitrarily}.

\begin{figure}[ht]
\begin{minipage}{0.45\linewidth}
\begin{ecode}
  let rec sumList xs =
    match xs with
    | []   -> (*@\hlTree{\hlSherrloc{[]}}@*)
    | h::t -> h (*@\hlTree{+}@*) (*@\hlFix{sumList t}@*)
\end{ecode}
\end{minipage}
\begin{minipage}{0.49\linewidth}
\begin{verbatim}
File "sumList.ml", line 4, characters 16-25:
  This expression has type 'a list
  but an expression was expected of type int
\end{verbatim}
\end{minipage}
\caption{(left) An ill-typed \ocaml program that should sum the elements of a
  list, with highlights indicating three possible blame assignments based on:
  %
  (1) the \hlFix{\ocaml} compiler;
  %
  (2) the \hlSherrloc{fix} made by the programmer; and
  %
  (3) \hlTree{minimizing} the number of edits required.
  % The \underline{underlined} expressions are equally valid
  % locations to blame. The expression blamed by the \ocaml compiler
  % is in \textbf{bold}.
  %
  % FIXME: This bolding is ambiguous, because ``let rec'', ``match'' and
  % ``with'' are also bolded (by \\ecode)! You need to find another way to
  % highlight what ocaml is yelling about.
  %
  (right) The error reported by \ocaml.}
\label{fig:sumList}
\end{figure}