### 2017-04-16 paper/oopsla17-submission/evaluation.tex
tweak

\mysection{Evaluation}
\label{sec:evaluation}
\input{data}

We have implemented our technique for localizing type errors for a
purely functional subset of \ocaml with polymorphic types and functions.
%
We seek to answer three questions in our evaluation:
%
\begin{itemize}
\item \textbf{Blame Accuracy}
  %
  How often does \toolname
  blame a \emph{correct}
  location for the type error?
  (\autoref{sec:quantitative})
  %
  % We compare our technique with a variety of off-the-shelf classifiers
  % and find that our top-ranked blame assignments have an accuracy of
  % 72\%, compared to a state-of-the-art 56\%.
  % For how many ill-typed programs can we accurately predict the source
  % of the error?
\item \textbf{Feature Utility}
  %
  Which program \emph{features are required}
  to localize errors?
   (\autoref{sec:feature-utility})
  % How much do the features described in \autoref{sec:features}
  % contribute to our predictions?
\item \textbf{Interpretability}
  %
  %% Do the models learned by \toolname
  %% correspond to our intuition about
  %% the real causes of errors?
  Are the models \emph{interpretable} using
  our intuition about the causes of type errors?
  (\autoref{sec:qualitative})
\end{itemize}
%
%\mypara{Summary of Results}
%
In the sequel we present our experimental
methodology \autoref{sec:methodology} and
then drill into how we evaluated each of
the questions above.
%
However, for the impatient reader, we begin
with a quick summary of our main results:
%
%
%\begin{itemize}
%
%\item \textbf

\mypara{1. Data beats Algorithms}
Our main result is that for type error
localization, data is indeed unreasonably
effective \citep{halevy09}.
%
\toolname's most sophisticated
\emph{neural network}
based classifier's top-1 error
location blames the correct
sub-term 74\% of the time
which, a good 18 points
higher than the state of
the art \sherrloc's 56\%.
%
However, even \toolname's simple
\emph{logistic regression} based
classifier is correct 60\% of the time,
\ie 4 points better than \sherrloc.
%
When the top-3 locations are considered,
\toolname is correct 91\% of the time.

% \item \textbf
\mypara{2. Slicing is Critical}
%
However, data is effective \emph{only}
when irrelevant sub-terms have been
sliced out of consideration.
%
In fact, perhaps our most surprising
result is that type-error-slicing and
local syntax alone yields
a neural classifier that is 10 points
better than \ocaml and on par with
\sherrloc.
%
That is once we focus classifiers on
slices, local syntactic features
alone perform as well as the
state-of-the-art.

%\item \textbf
\mypara{3. Size doesn't matter, types do}
%
We find that (after slices)
contextual typing features
provide the biggest
improvement in accuracy.
%
Furthermore we find contextual syntax
features to be mostly (but not entirely)
redundant with typing features,
which supports the hypothesis that
the context's \emph{type} nicely
summarizes the properties of the
surrounding expressions.
%
Finally, we found that the \emph{size}
of the sub-expression was not very useful.
This was unexpected, as we thought
smaller expressions would be simpler, and
hence, more likely causes.

% \item \textbf
\mypara{4. Models Learn Typing Rules}
%
Finally, by investigating a few of the
predictions made by the \emph{decision tree}-based
models, we found that the models
do indeed capture some simple and intuitive
rules for predicting well-typedness.
%
For example, if the left child of an application
is a function, then the application is likely
correct.

% in an application, if the
% left argument is a function then the
% error is likely on the right sub-term;
% in a function definition
% \RJ{fixme: orig
% defining a function is a fine thing
% to do?}
%
%\end{itemize}
% \RJ{eric please check the numbers -- we say 71, 72, 74, 91, 92, 94?}

\mysubsection{Methodology}
\label{sec:methodology}

We answer our questions on two sets of data gathered from the
undergraduate Programming Languages course at
\begin{anonsuppress}
UC San Diego (IRB \#140608).
\end{anonsuppress}
\begin{noanonsuppress}
AUTHOR's INSTITUTION.
\end{noanonsuppress}
%
We recorded each interaction with the \ocaml top-level system over the
course of the first three assignments, capturing
ill-typed programs and, crucially, their subsequent fixes.
%
The first dataset comes from the Spring 2014 class (\SPRING), with a
cohort of 46 students. The second comes from the Fall 2015 class
(\FALL), with a cohort of 56 students.
%
The extracted programs are relatively small, but they demonstrate a
range of functional programming idioms, \eg higher-order functions and
(polymorphic) algebraic data types.

\mypara{Feature Selection}
We extract a set of 277 features from each sub-expression in a student
program, including:
%
\begin{enumerate}
\item 44 local syntactic features. In addition to the syntax of \lang,
  we support the full range of arithmetic operators (integer and
  floating point), equality and comparison operators, character and
  string literals, and a user-defined % |expr| type of simple
  arithmetic
  expressions. We discuss the challenge of supporting other
  % user-defined
  types in \autoref{sec:discussion}.
\item 176 contextual syntactic features. For each sub-expression we
  additionally extract the local syntactic features of its parent and
  first, second, and third (left-to-right) children. If an expression
  does not have a parent or children, these features will simply be
  disabled. If an expression has more than three children, the
  classifiers will receive no information about the additional
  children.
\item 55 typing features. In addition to the types of \lang, we support
  |int|s, |float|s, |char|s, |string|s, and the user-defined |expr|
  mentioned above. These features are extracted for each sub-expression
  and its context. % for the contextual sub-expressions.
\item One feature denoting the size of each sub-expression.
\item One feature denoting whether each sub-expression is part of the
  minimal type error slice. We use this feature as a ``hard''
  constraint, sub-expressions that are not part of the minimal slice
  will be preemptively discarded. We justify this decision in
  \autoref{sec:feature-utility}.
\end{enumerate}

\mypara{Blame Oracle}
Recall from \autoref{sec:labels} that we automatically extract a blame
oracle for each ill-typed program from the (AST) diff between it and the
student's eventual fix.
%
A disadvantage of using diffs in this manner is that students may have
made many, potentially unrelated, changes between compilations; at some
point the ``fix'' becomes a ``rewrite''.
%
We do not wish to consider the ``rewrites'' in our evaluation, so we
discard outliers where the fraction of expressions that have changed is
more than one standard deviation above the mean, establishing a diff
threshold of 44\%.
%
This accounts for roughly 14\% of each dataset, leaving us with
2,425 program pairs for \SPRING and 2,325 pairs for \FALL.

% we discard any program pairs where more than 40\%
% of the sub-expressions have changed.
% %
% We picked 40\% as an estimate of the inflection point where we could
% still retain the large majority of program pairs.
% % FIXME: Can you say that this dataset curation is similar to any other
% % datasets (e.g., the washington one)? Anything you could cite and discuss
% % here would take some of the pressure off.

\mypara{Accuracy Metric}
All of the tools we compare (with the exception of the standard \ocaml
compiler) can produce a list of potential error locations.
%
However, in a study of fault localization techniques,
\citet{Kochhar2016-oc} show that most developers will not consider more
than around five potential error locations before falling back to manual
debugging.
%
Type errors are relatively simple in comparison to general fault
localization, thus we limit our evaluation to the top three predictions
of each tool.
%
We evaluate each tool on whether a changed expression occurred in its
top one, top two, or top three predictions.

\input{evaluation-accuracy}
\input{evaluation-utility}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: