### 2017-07-24 paper/oopsla17-cameraready/data.tex
copy into oopsla17-cameraready

\pgfplotstableset{col sep=comma}

% \pgfplotstableread{../../data/sp14/op+type+size/linear/results.csv}{\FeatureLinearBench}
% \pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context+type+size/linear/results.csv}
% \pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context-has+type+size/linear/results.csv}
% \pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context-count+type+size/linear/results.csv}
% \pgfplotstableread{../../data/sp14/op+type+size/hidden-10/results.csv}{\FeatureHiddenTBench}
% \pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context+type+size/hidden-10/results.csv}
% \pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context-has+type+size/hidden-10/results.csv}
% \pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context-count+type+size/hidden-10/results.csv}
% \pgfplotstableread{../../data/sp14/op+type+size/hidden-500/results.csv}{\FeatureHiddenFHBench}
% \pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context+type+size/hidden-500/results.csv}
% \pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context-has+type+size/hidden-500/results.csv}
% \pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}

% \pgfplotstableread{../../data/sp14/op+type+size/hidden-10/results.csv}{\HiddenBench}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-25/results.csv}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-50/results.csv}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-100/results.csv}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-250/results.csv}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-500/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-10/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-25/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-50/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-100/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-250/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}

\pgfplotstableread{../../data/sp14/baseline.csv}{\SpringBench}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/ocaml/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/mycroft/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/sherrloc/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/linear/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/decision-tree/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/random-forest/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/hidden-10/results.csv}
\pgfplotstablevertcat{\SpringBench}{../../data/sp14/op+context+type+size/hidden-500/results.csv}

\pgfplotstableread{../../data/fa15/baseline.csv}{\FallBench}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/ocaml/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/mycroft/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/sherrloc/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/linear/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/decision-tree/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/random-forest/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/hidden-10/results.csv}
\pgfplotstablevertcat{\FallBench}{../../data/fa15/op+context+type+size/hidden-500/results.csv}

\pgfplotstableread{../../models/linear-op+slice-no-slice.cross.csv}{\SliceLinearBench}
\pgfplotstablevertcat{\SliceLinearBench}{../../models/linear-op+slice.cross.csv}
\pgfplotstablevertcat{\SliceLinearBench}{../../models/linear-op+slice-only-slice.cross.csv}
\pgfplotstableread{../../models/hidden-500-op+slice-no-slice.cross.csv}{\SliceHiddenBench}
\pgfplotstablevertcat{\SliceHiddenBench}{../../models/hidden-500-op+slice.cross.csv}
\pgfplotstablevertcat{\SliceHiddenBench}{../../models/hidden-500-op+slice-only-slice.cross.csv}
% \pgfplotstablecreatecol[create col/assign/.code={%
%     \edef\entry{\thisrow{features}/\thisrow{model}}
%     \pgfkeyslet{/pgfplots/table/create col/next content}\entry
%   }]{tool}{\SliceBench}

\pgfplotstableread{../../models/linear-op.cross.csv}{\FeatureLinearBench}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+size.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+context.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+type.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+context+size.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+type+size.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+context+type.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{../../models/linear-op+context+type+size.cross.csv}
\pgfplotstableread{../../models/linear-op.cross.csv}{\FeatureHiddenBench}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+size.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+context.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+type.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+context+size.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+type+size.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+context+type.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{../../models/hidden-500-op+context+type+size.cross.csv}
% \pgfplotstablecreatecol[create col/assign/.code={%
%     \edef\entry{\thisrow{features}/\thisrow{model}}
%     \pgfkeyslet{/pgfplots/table/create col/next content}\entry
%   }]{tool}{\FeatureBench}