### 2017-07-26 paper/oopsla17-cameraready/evaluation-user-study-graph.tex
fiddle with graphs

\begin{figure}[t]
\centering
%% EXPLAIN
\begin{tikzpicture}
\begin{axis}[
  ybar,
  % width=6cm,
  height=6cm,
  title={\large Explanation},
  ylabel={\% Correct},
  bar width=0.5cm,
  ymin=0,
  ymax=1,
  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
  ytick style={draw=none},
  ymajorgrids = true,
  enlarge x limits=0.25,
  symbolic x coords={{sepConcat}, {padZero}, {mulByDigit}},
  xtick=data,
  xtick style={draw=none},
  xticklabel style={align=center},
  xticklabels={
    {\texttt{sepConcat}\\($p=0.48$)},
    {\texttt{padZero}\\($p=0.097$)},
    {\texttt{mulByDigit}\\($p=0.083$)}
  },
  % x tick label style={rotate=45, anchor=north east},
  %x tick label style={font=\small},
  y tick label style={font=\small},
  reverse legend,
  transpose legend,
  legend style={legend pos = south west, font=\footnotesize},
]

% ES: NOTE: ORDER OF PLOTS/LEGEND ENTRIES MATTERS

\addplot[draw=black, fill=green2, bar shift=.25cm, error bars/.cd, y dir=both, y explicit] coordinates {
  (sepConcat, 0.8125)  +- (0, 0.0976)
  (padZero, 1.00)      +- (0, 0)
  (mulByDigit, 0.8125) +- (0, 0.0976)
};
\addlegendentry{\toolname}

\addplot[draw=black, fill=blue2, bar shift=-.25cm, error bars/.cd, y dir=both, y explicit] coordinates {
  (sepConcat, 0.80)    +- (0, 0.1033)
  (padZero, 0.875)     +- (0, 0.0827)
  (mulByDigit, 0.5714) +- (0, 0.1323)
};
\addlegendentry{\sherrloc}
\end{axis}
\end{tikzpicture}
%% FIX
\begin{tikzpicture}
\begin{axis}[
  ybar,
  % width=6cm,
  height=6cm,
  title={\large Fix},
  ylabel={\% Correct},
  bar width=0.5cm,
  ymin=0,
  ymax=1,
  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
  ytick style={draw=none},
  ymajorgrids = true,
  enlarge x limits=0.25,
  symbolic x coords={{sepConcat}, {padZero}, {mulByDigit}},
  xtick=data,
  xtick style={draw=none},
  xticklabel style={align=center},
  xticklabels={
    {\texttt{sepConcat}\\($p=0.57$)},
    {\texttt{padZero}\\($p=0.33$)},
    {\texttt{mulByDigit}\\($p=0.31$)}
  },
  % x tick label style={rotate=45, anchor=north east},
  %x tick label style={font=\small},
  y tick label style={font=\small},
  reverse legend,
  transpose legend,
  legend style={legend pos = south west, font=\footnotesize},
]

% ES: NOTE: ORDER OF PLOTS/LEGEND ENTRIES MATTERS

\addplot[draw=black, fill=green2, bar shift=.25cm, error bars/.cd, y dir=both, y explicit] coordinates {
  (sepConcat, 0.8125)  +- (0, 0.0976)
  (padZero, 0.9286)    +- (0, 0.0688)
  (mulByDigit, 0.7143) +- (0, 0.1207)
};
\addlegendentry{\toolname}

\addplot[draw=black, fill=blue2, bar shift=-.25cm, error bars/.cd, y dir=both, y explicit] coordinates {
  (sepConcat, 0.83)    +- (0, 0.1076)
  (padZero, 0.875)     +- (0, 0.0827)
  (mulByDigit, 0.6154) +- (0, 0.1349)
};
\addlegendentry{\sherrloc}
\end{axis}
\end{tikzpicture}
\caption[A classification of students' explanations and fixes for type
  errors, given either \sherrloc or \toolname's blame assignment.]
  {A classification of students' explanations and fixes for type
  errors, given either \sherrloc or \toolname's blame assignment.
  %
  The students given \toolname's location generally scored
  better than those given \sherrloc's.
  %
  We report the result of a one-sided Mann-Whitney $U$ test for
  statistical significance in parentheses.}
\label{fig:user-study-results}
\end{figure}