### 2017-08-24 paper/oopsla17-cameraready/data.tex
checkpoint

\pgfplotstableset{col sep=comma}

% \pgfplotstableread{../../data/sp14/op+type+size/linear/results.csv}{\FeatureLinearBench}
% \pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context+type+size/linear/results.csv}
% \pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context-has+type+size/linear/results.csv}
% \pgfplotstablevertcat{\FeatureLinearBench}{../../data/sp14/op+context-count+type+size/linear/results.csv}
% \pgfplotstableread{../../data/sp14/op+type+size/hidden-10/results.csv}{\FeatureHiddenTBench}
% \pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context+type+size/hidden-10/results.csv}
% \pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context-has+type+size/hidden-10/results.csv}
% \pgfplotstablevertcat{\FeatureHiddenTBench}{../../data/sp14/op+context-count+type+size/hidden-10/results.csv}
% \pgfplotstableread{../../data/sp14/op+type+size/hidden-500/results.csv}{\FeatureHiddenFHBench}
% \pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context+type+size/hidden-500/results.csv}
% \pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context-has+type+size/hidden-500/results.csv}
% \pgfplotstablevertcat{\FeatureHiddenFHBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}

% \pgfplotstableread{../../data/sp14/op+type+size/hidden-10/results.csv}{\HiddenBench}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-25/results.csv}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-50/results.csv}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-100/results.csv}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-250/results.csv}
% \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+type+size/hidden-500/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-10/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-25/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-50/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-100/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-250/results.csv}
% % \pgfplotstablevertcat{\HiddenBench}{../../data/sp14/op+context-count+type+size/hidden-500/results.csv}

\pgfplotstableread{data/sp14-baseline.csv}{\SpringBench}
\pgfplotstablevertcat{\SpringBench}{data/sp14-ocaml-results.csv}
\pgfplotstablevertcat{\SpringBench}{data/sp14-mycroft-results.csv}
\pgfplotstablevertcat{\SpringBench}{data/sp14-sherrloc-results.csv}
\pgfplotstablevertcat{\SpringBench}{data/sp14-op+context+type+size-linear-results.csv}
\pgfplotstablevertcat{\SpringBench}{data/sp14-op+context+type+size-decision-tree-results.csv}
\pgfplotstablevertcat{\SpringBench}{data/sp14-op+context+type+size-random-forest-results.csv}
\pgfplotstablevertcat{\SpringBench}{data/sp14-op+context+type+size-hidden-10-results.csv}
\pgfplotstablevertcat{\SpringBench}{data/sp14-op+context+type+size-hidden-500-results.csv}

\pgfplotstableread{data/fa15-baseline.csv}{\FallBench}
\pgfplotstablevertcat{\FallBench}{data/fa15-ocaml-results.csv}
\pgfplotstablevertcat{\FallBench}{data/fa15-mycroft-results.csv}
\pgfplotstablevertcat{\FallBench}{data/fa15-sherrloc-results.csv}
\pgfplotstablevertcat{\FallBench}{data/fa15-op+context+type+size-linear-results.csv}
\pgfplotstablevertcat{\FallBench}{data/fa15-op+context+type+size-decision-tree-results.csv}
\pgfplotstablevertcat{\FallBench}{data/fa15-op+context+type+size-random-forest-results.csv}
\pgfplotstablevertcat{\FallBench}{data/fa15-op+context+type+size-hidden-10-results.csv}
\pgfplotstablevertcat{\FallBench}{data/fa15-op+context+type+size-hidden-500-results.csv}

\pgfplotstableread{models/linear-op+slice-no-slice.cross.csv}{\SliceLinearBench}
\pgfplotstablevertcat{\SliceLinearBench}{models/linear-op+slice.cross.csv}
\pgfplotstablevertcat{\SliceLinearBench}{models/linear-op+slice-only-slice.cross.csv}
\pgfplotstableread{models/hidden-500-op+slice-no-slice.cross.csv}{\SliceHiddenBench}
\pgfplotstablevertcat{\SliceHiddenBench}{models/hidden-500-op+slice.cross.csv}
\pgfplotstablevertcat{\SliceHiddenBench}{models/hidden-500-op+slice-only-slice.cross.csv}
% \pgfplotstablecreatecol[create col/assign/.code={%
%     \edef\entry{\thisrow{features}/\thisrow{model}}
%     \pgfkeyslet{/pgfplots/table/create col/next content}\entry
%   }]{tool}{\SliceBench}

\pgfplotstableread{models/linear-op.cross.csv}{\FeatureLinearBench}
\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+size.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+context.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+type.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+context+size.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+type+size.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+context+type.cross.csv}
\pgfplotstablevertcat{\FeatureLinearBench}{models/linear-op+context+type+size.cross.csv}
\pgfplotstableread{models/linear-op.cross.csv}{\FeatureHiddenBench}
\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+size.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+context.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+type.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+context+size.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+type+size.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+context+type.cross.csv}
\pgfplotstablevertcat{\FeatureHiddenBench}{models/hidden-500-op+context+type+size.cross.csv}
% \pgfplotstablecreatecol[create col/assign/.code={%
%     \edef\entry{\thisrow{features}/\thisrow{model}}
%     \pgfkeyslet{/pgfplots/table/create col/next content}\entry
%   }]{tool}{\FeatureBench}